---
title: "Class Assignment: MPG tables and charts"
author: "Yunyi Zhang"
date: "September 16, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)                ## loading the entire knitr package
library(tidyverse)              ####### change# load tidyverse
library(esquisse)             ## use esquisse to setup the basic plots

library(kableExtra)
library(magrittr)
opts_chunk$set(echo = FALSE)
```

# Using the tidyverse (especially dplyr) for MPG

Recall assignment:

Tables should include a comparison of city and highway mileage by the class of car and the best three car models for city and highway mileage for all the years in which data is available.  

Plot the data displaying as much of the data as you can. Put continuous variables on the axes and include a locally smoothed regression line to show the relationship with mileage.  Also make a box plot comparing city and highway MPG by class of car.  


## Tables




```{r}

data(mpg)     ## load the data


mpg$cyl <- as.factor(mpg$cyl)   ## convert discrete variables to factors so that they plot
mpg$drv <- as.factor(mpg$drv)   ## as different colors, not gradations of one color

mpg$cty <- as.double(mpg$cty)

mpg_1 <- select(mpg, cty, hwy, class )


mpg_1 %<>% group_by(class) %>% 
  summarise(City=mean(cty),
            Highway=mean(hwy)) %>% 
  arrange(class, City, Highway) 








                                
class <- unique(mpg$class)      ## define a vector for car classes

cty_mean <- rep(0, length(class))  ## define variables to hold mean mileage
hwy_mean <- rep(0,length(class))   ## for each class of car

for(i in 1:length(class)){                          ## calculate the mean city and 
  cty_mean[i] = mean(mpg$cty[mpg$class==class[i]])  ## highway mileage for each 
  hwy_mean[i] = mean(mpg$hwy[mpg$class==class[i]])  ## class of car
}

cty_mean <- round(cty_mean, 2)                      ## round the results to 2 decimal pla
hwy_mean <- round(hwy_mean, 2)

tbl_1 <- cbind(class, cty_mean,hwy_mean)      ## form an array for the table

kable(tbl_1, format = "latex", booktabs=TRUE, digits = 2,     ## call kable to make the table
      col.names = c("Class", "City", "Highway"), 
      caption = "Mean City and Highway MPG by Car Class" )      






```


The next table is a bit tricky. 



```{r}

##  You probably want to get started with a little investigation. 
##  Using unique(mpg$year), you find out that there are only two years - 1999, 2008.


options(digits = 3)
data <-select(mpg,model,year,cty,hwy)
mod_08_cty_mean <- mpg %>% dplyr::filter(year == 2008) %>%
group_by(model) %>% dplyr::summarise(City= mean(cty)) %>% arrange(desc(City)) %>%   head(3)
mod_99_cty_mean <- mpg %>% dplyr::filter(year == 1999) %>%
group_by(model) %>% dplyr::summarise(City= mean(cty)) %>% arrange(desc(City)) %>%   head(3)
mod_08_hwy_mean <- mpg %>% dplyr::filter(year == 2008) %>%
group_by(model) %>% dplyr::summarise(City= mean(hwy)) %>% arrange(desc(City)) %>%   head(3)
mod_99_hwy_mean <- mpg %>% dplyr::filter(year == 1999) %>%
group_by(model) %>% dplyr::summarise(City= mean(hwy)) %>% arrange(desc(City)) %>%   head(3)

table2 <-cbind(mod_99_cty_mean, mod_99_hwy_mean, mod_08_cty_mean, mod_08_hwy_mean)

colnames(table2) <- c('Model', 'Milage', "Model", "Milage",'Model', 'Milage',"Model", "Milage")
kable(table2, digits = 2, format = "latex", booktabs=TRUE, ,caption = "Top 3 MPG Performing Cars: 1999, 2008") %>% kable_styling() %>%
  add_header_above(c("City 1999"=2, 
                     "Highway 1999"=2,
                     "City 2008"=2,
                   "Highway 2008"=2))

```




```{r}





library(ggplot2)

ggplot(mpg) +
 aes(x = displ, y = cty) +
 geom_point(aes(color=class),size=2) + geom_smooth() +
 scale_color_hue() +
 theme_minimal(base_size = 16) +
 facet_wrap(vars(year)) + 
  labs(x = "Engine Displacement",y = "City Mileage" ,title = "City MPG by Class of Car: 1999, 2008")
 
 
 
 ggplot(mpg) +
 aes(x = displ, y = hwy) +
 geom_point(aes(color=class),size=2) + geom_smooth() +
 scale_color_hue() +
 theme_minimal(base_size = 16) +
 facet_wrap(vars(year)) + 
  labs(x = "Engine Displacement", y = "Highway Mileage", title = "Highway MPG by Class of Car: 1999, 2008")
 
 

##esquisser(mpg)

```

Boxplots

```{r}



ggplot(mpg) +
 aes(x = class, y = cty) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal(base_size=16) +
  labs(x = "Car Class", y = "City Mileage", title = "City MPG by Class of Car: 1999, 2008")


ggplot(mpg) +
 aes(x = class, y = hwy) +
 geom_boxplot(fill = "#0c4c8a") +
 theme_minimal(base_size=16) +
  labs(x = "Car Class", y = "Highway Mileage", title = "Highway MPG by Class of Car: 1999, 2008")





```
1234

